Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-04-05 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Authors

% IEAs
Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
whose fitness evaluations are performed by humans through an interactive 
system \cite{eiben2015interactive}.
They are usually applied in problems where the fitness function is not known or simply
does not exist, and the result of optimization should fit a certain human need or
desire such as an aesthetic ideal.  
That is why their use cases include the evolution of objects with subjective characteristics,
such as visual appeal and attractiveness \cite{biomorphs} as well as others where human behavior is 
considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
or creativity \cite{yu2011cooks}. In the cases when 
human interaction is responsible of other 
aspects of the evolutionary process, these IEC methods are classified by some authors  
as human-based evolutionary computation \cite{972056} 
or as human-based computation \cite{quinn2011human}.

% Human Based Computation
IEC systems are an interesting venue of research, since they have demonstrated 
their ability for effectively 
producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
as well as other types of artifacts in many other domains \cite{ie1}. 
% Problems
However, the necessary intervention of humans brings certain challenges 
to designers of IEC methods; namely, human evaluations are scarce, slow and expensive, there is
human fatigue caused by the interaction \cite{ie1}, and also boredom arises
when users evaluate a large number of phenotypes, 
many of which are not interesting or are very similar to each other.
% C-IEAs
Moreover, the performance of these systems effectively depends on the number of users
they are able to include; to reach more users,
IEC systems are some times developed as web applications depending on visitors to help
with the search, using both anonymous and registered users. Some systems 
employ a collaborative technique, where several users participate in 
the evaluation, this method is called Collaborative-IEC (C-IEC)
\cite{picbreeder,seyama2016development,wagy2014collective}.
% Volunteers 
Including an C-IEC in a volunteer system can lower the requirements for
participants in the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
volunteer's lack of accountability, and the need to build trust between participants and project
owners. Other issues of interest for project owners are also the difficulty of predicting 
the amount of time and resources a volunteer is willing to spend on the system, 
and how they decide if they participate or not \cite{JJ:2016}.

@@

For example, in a C-IEC application fitness assignment depends on the
actions of a social network of users.  These actions are triggered when 
they tag, share, rate, store or delete a phenotype. 
Then, the selection of parents could depend on the previous actions, leveraging information, 
such as the fact that they both have the same tag, or were shared by
similar users.

@@

It was shown how the estimation of fitness of candidate solutions
can take advantage of a graph model representation. Also, the graph
offers another view of the system that it would have been
difficult to replicate with a relational database. 
% More conclusions and some way of relating the three case studies
% would be something that is missing and should definitely go here - JJ
% No longer needed as they where removed - Mario

>>>>

% IEAs
Interactive evolutionary computation (IEC) systems are, in general, evolutionary methods
whose fitness evaluations are performed by humans through an interactive 
system \cite{eiben2015interactive}.
% They are usually applied in problems where the fitness function is not known or simply
% does not exist, and the result of optimization should fit a certain human need or
% desire such as an aesthetic ideal.  
% That is why their use cases include the evolution of objects with subjective characteristics,
% such as visual appeal and attractiveness \cite{biomorphs} as well as others where human behavior is 
% considered, for instance the optimization of teamwork \cite{kosorukoff2002evolutionary}
% or creativity \cite{yu2011cooks}. In the cases when 
% human interaction is responsible of other 
% aspects of the evolutionary process, these IEC methods are classified by some authors  
% as human-based evolutionary computation \cite{972056} 
% or as human-based computation \cite{quinn2011human}.

% Human Based Computation
% IEC systems are an interesting venue of research, since they have demonstrated 
% their ability for effectively 
% producing art and design \cite{Bentley:1999:intro,Sims:1991,todd:1992,evoeco},
% as well as other types of artifacts in many other domains \cite{ie1}. 
% Problems
However, the necessary intervention of humans brings certain challenges 
to designers of IEC methods; namely, human evaluations are scarce, slow and expensive, there is
human fatigue caused by the interaction \cite{ie1}, and also boredom arises
when users evaluate a large number of phenotypes, 
many of which are not interesting or are very similar to each other.
% % C-IEAs
% Moreover, the performance of these systems effectively depends on the number of users
% they are able to include; to reach more users,
% IEC systems are some times developed as web applications depending on visitors to help
% with the search, using both anonymous and registered users. Some systems 
% employ a collaborative technique, where several users participate in 
% the evaluation, this method is called Collaborative-IEC (C-IEC)
% \cite{picbreeder,seyama2016development,wagy2014collective}.
% Volunteers 
Including an C-IEC in a volunteer system can lower the requirements for
participants in the experiment thus increasing the {\em performance} of the whole system. But using a volunteer based 
system raises other issues \cite{sarmenta2001volunteer,web:BOINC}, such as the 
volunteer's lack of accountability, and the need to build trust between participants and project
owners. Other issues of interest for project owners are also the difficulty of predicting 
the amount of time and resources a volunteer is willing to spend on the system, 
and how they decide if they participate or not \cite{JJ:2016}.

@@

% For example, in a C-IEC application fitness assignment depends on the
% actions of a social network of users.  These actions are triggered when 
% they tag, share, rate, store or delete a phenotype. 
% Then, the selection of parents could depend on the previous actions, leveraging information, 
% such as the fact that they both have the same tag, or were shared by
% similar users.

@@

### 2017-04-05 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-front.tex
Authors

\begin{document}
\title{Modeling Human Interactions in Collaborative Interactive 
 Evolutionary Computation}
%\titlenote{Produces the permission block, and
%  copyright information}
\subtitle{Anonymized Version}
%\subtitlenote{The full version of the author's guide is available as
%  \texttt{acmart.pdf} document}

\author{Ben Trovato}
%\authornote{Dr.~Trovato insisted his name be first.}
\orcid{1234-5678-9012}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin} 
  \state{Ohio} 
  \postcode{43017-6221}
}
\email{trovato@corporation.com}

\author{G.K.M. Tobin}
%\authornote{The secretary disavows any knowledge of this author's actions.}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin} 
  \state{Ohio} 
  \postcode{43017-6221}
}
\email{webmaster@marysville-ohio.com}

\author{Lars Th{\o}rv{\"a}ld}
%\authornote{This author is the
%  one who did all the really hard work.}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla} 
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Lawrence P. Leipuner}
\affiliation{
  \institution{Brookhaven Laboratories}
  \streetaddress{P.O. Box 5000}}
\email{lleipuner@researchlabs.org}

\begin{abstract}
The necessary intervention of humans in interactive evolutionary
computational systems has inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of artifacts.
To tackle these issues, in this paper we
propose a human-centered framework to model complex interactions on these systems.
A case study is presented where the model is applied in the 
development of a collaborative evolutionary interactive system. 
Both conceptual and implementation details are provided where the   
technique is used to measure and increase user engagement and participation.
Our experiments show that the model can be successfully applied in a
gamification technique developed to increase user engagement, which
implies that this technique can successfully be used to 
decrease user fatigue and boredom, and thus increase the {\em performance} of the
interactive system.
\end{abstract}

>>>>

\begin{document}
\title{Modeling Human Interactions in Collaborative Interactive 
 Evolutionary Computation}

\author{Mario Garc\'ia-Valdez,\\ Jos\'e-C  Romero}
\affiliation{%
  \institution{Dept. of Graduate Studies,\\ Instituto
Tecnol\'ogico de Tijuana, Tijuana, M\'exico}
  }
\email{(mario|jcromero)tectijuana.edu.mx}

\author{Juan-J.~Merelo}
\affiliation{%
  \institution{Dept. of Computer Architecture and Technology, University of Granada, Granada, Spain}
  }
\email{jmerelo@ugr.es}

\author{Francisco Fern\'andez-de-Vega}
\affiliation{%
  \institution{Grupo de Evoluci\'on Artificial,\\
Universidad de Extremadura, Spain}
  }
\email{fcofdez@unex.es}

\begin{abstract}
The necessary intervention of humans in interactive evolutionary
computational systems has inherent drawbacks arising from the very nature of 
the algorithms, namely, the human fatigue caused by the interaction, and
the boredom arising when users evaluate a large number of artifacts.
To tackle these issues, in this paper we
propose a human-centered framework to model complex interactions on these systems.
A case study is presented where the model is applied in the 
development of a collaborative evolutionary interactive system. 
% Both conceptual and implementation details are provided where the   
% technique is used to measure and increase user engagement and participation.
% Our experiments show that the model can be successfully applied in a
% gamification technique developed to increase user engagement, which
% implies that this technique can successfully be used to 
% decrease user fatigue and boredom, and thus increase the {\em performance} of the
% interactive system.
\end{abstract}
